---
title: "Proeftentamen_heights_weights"
author: "Marc Teunis / Chris van Oevelen"
date: '`r Sys.Date()`'
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(forcats)
library(here)
```

```{r importing, include=FALSE}
height_weight <-read_csv(here("data/proeftentamen/height_weight.csv"))
```

##### __Inspection of datafile:__
###### __Question1:__ Show the first two lines of the datafile in a table
```{r inspection_head, echo=FALSE}
knitr::kable(
head(height_weight, n=1),
caption = "Table1: The dataset contains heights, weights and 8 more variables about 4000 people. "
)
```

##### __Data_wrangling. Creation of a new variable: body mass index (bmi):__ 
###### __Question2:__ Calculate the body mass index (bmi) of each person and store in new variable named "bmi". Bmi = weight (in kg) / height (in meters) ^2  

The unit of weight = lbs and of height = inches. To create body mass index the variable weight and height needs to be converted to respectively kilogram (kg) and meters (m):  
1 inch = 2.54 cm  
1 lbs = 0.45359237 kg  

```{r conversion, echo = FALSE}
height_weight <-height_weight %>% 
  mutate(height_m = height * 0.0254 ,
         weight_kg = weight * 0.45359237 , 
         bmi=weight_kg / (height_m^2))
height_weight_slc <- height_weight %>% select(income:education, bmi)
```

###### __Question3:__ Make a selection of variables and show the first two lines of the selected variables in a table. Variables to be selected are: income, height, weight, sex, race, education and bmi

```{r table2, echo = FALSE}
knitr::kable( 
  head(height_weight_slc, n=1), 
  caption = "Table2: Selection of variables of the datafile"      
)
```

##### __Data analysis: data exploration__  
###### __Question4:__ Which race has the highest % of persons with a weight > 200 lbs?  

```{r 200_lbs, echo=FALSE}
height_weight_slc_na <-height_weight_slc %>% filter(!is.na(weight)) 
knitr::kable(
height_weight_slc_na %>% group_by(race) %>% 
  summarize(race_200 = sum(weight > 200, na.rm=T),
            race_n = n(),
            fraction_more_than_200lbs = round(race_200 / race_n, digits = 3) ) %>% 
  select(race, fraction_more_than_200lbs) ,
caption = "Table3: Black race has highest fraction of weight > 200" )
```

###### __Question5:__ Filter the dataset to extract the top 5 tallest hispanic males in the dataset.

```{r tallest_hispanic, echo=FALSE}
knitr::kable(
  tallest_hispanic_men <- height_weight_slc %>%
  filter(race == "hispanic" & sex == "male") %>%
  arrange(desc(height)) %>% head(n=5),
  caption = "Table4: Tallest Hispanic men" )
```

###### __Question6:__ Find the two tallest women in the data. 

```{r tallest_women, echo = FALSE}
knitr::kable(
  rank_women_2 <- height_weight_slc %>%
  filter(sex == "female") %>%
  arrange(desc(height)) %>% head(n=2),
  caption = "Table5: Tallest woman" )
```

###### __Question7:__ Find all men that match the height and weight of the 2nd tallest woman of question 6  
```{r match, echo = FALSE}
knitr::kable(
  match_1 <- height_weight_slc %>%
  filter(sex == "male" ,  
         height == rank_women_2$height[2],
         weight == rank_women_2$weight[2] ),
  caption = "Table6: List of men matching the 2nd tallest woman by weight and height"
)
```

##### __Data analysis: difference between bmi values per race and sex__
###### __Question8:__ Make a summary of the bmi values for each group (race x sex = 4 x 2 = 8 groups)

```{r bmi_summary, echo=FALSE}
knitr::kable(
  height_weight_slc %>% 
    group_by (race, sex ) %>% 
    summarize (mean_bmi = mean(bmi, na.rm=TRUE),
               sd_bmi = sd(bmi, na.rm=TRUE),
               median_bmi = median(bmi, na.rm=TRUE),
               min = min (bmi, na.rm = TRUE),
               max = max (bmi, na.rm = TRUE),
               count=n()),
  caption = "Table7: Summary of the bmi values per race and sex"
)
```

###### __Question9:__ Visualise the bmi data for the 8 groups by a boxplot

```{r boxplot, echo=FALSE, warning=FALSE}
height_weight_slc %>% ggplot(aes(x=race, y=bmi, fill = sex )) + 
  geom_boxplot() 
```

###### __Question10:__ Perform Shapiro-wilk test for each of the 8 groups

H~0~: Data of each group is normally distributed  
H~1~: Data of each group is not normally distributed

```{r Shapiro_wilk, echo = FALSE}
knitr::kable(
  height_weight_slc %>% 
    group_by(race, sex) %>% 
    summarize (SW = shapiro.test(bmi)$p.value),
    caption = "Table8: P-values of SW test per race and sex"
)
```

Conclusion: None have the race groups have both p-values of the SW test > 0.05. Use a non-parametric test for further analysis. 

###### __Question11:__ Perform the right statistical test (based on the outcome of the Shapiro-Wilk test

Based on the SW test we will use a unpaired wilcoxon test  
H~0~: There is no statistically significant difference in bmi between male and female per race
H~1~: There is statistically significant difference in bmi between male and female per race

```{r wilcoxon, echo=FALSE, warning=FALSE}
knitr::kable(
  height_weight_slc %>% group_by(race) %>% 
  summarize (wx = wilcox.test(formula = bmi ~ sex, 
                              paired = FALSE)$p.value),
  caption = "Table9: P-values of wilcoxon test for each race"
)
```

Conclusion: the groups black, hispanic and other have p-value < 0.05. Reject H~0~ and accept H~1~: There is a statistically significant difference between male and female. For the multiracial group p-value > 0.05. Accept H~0~. No difference. __Note:__ this group has relative small number of data values (female: n=21, male: n=26)

#### __Data analysis: correlation between education and income__

###### __Question12:__  Filter row with education level > 5

```{r filter_education, include=FALSE}
height_weight_slc_flt <- height_weight_slc %>% 
  filter(education > 5)
```

###### __Question13:__ Inspect data by making 8 scatter plots. Each graph in a separate panel  

```{r line_graphs, echo=FALSE}
height_weight_slc_flt %>% ggplot(aes(education, income)) +
  geom_jitter() +
  geom_smooth(aes(colour = race), method = "lm") +
  facet_grid(sex~race)
```

###### __Question14:__ Determine the R squared of each regression line for the 8 groups  

```{r R-squared, echo=FALSE}
group_race_sex <-height_weight_slc_flt %>% 
  group_split(height_weight_slc_flt$race, 
              height_weight_slc_flt$sex ) 

lm_rsquared <-group_race_sex %>% map(lm, formula = income ~ education) %>% 
  map(summary) %>% 
  map_dbl("adj.r.squared") %>% round (digits = 3)

count_race_sex <- height_weight_slc_flt %>% 
  count(race,sex)

knitr::kable(
  tibble(race = count_race_sex$race, 
       sex = count_race_sex$sex, 
       count = count_race_sex$n, 
       r_squared=lm_rsquared),
      caption = "Table10: R-squared values of regression lines per race and sex"
)
```

Conclusion: R squared values are generally low, except 0,394 of group multiracial::female. However, the analysis is based on n=5. Generally there is a absent or weak correation between education and income based on this dataset
